Global patterns of forest autotrophic carbon fluxes

### 2020-08-06 manuscript/manuscript-draft.tex
working on refs... seem to have broken it :(

The latitudinal gradient in forest C flux rates, along with altitudinal
gradients \citep{girardin_net_2010, malhi_variation_2017}, is driven
primarily by climate, which is a significant driver of C fluxes across
broad spatial scales
\citep{luyssaert_co_2007, cleveland_relationships_2011, hursh_sensitivity_2017}.
However, there is little consensus as to the shapes of these
relationships or the best predictor variables. The majority of studies
have focused on exploring the relationships of C fluxes to mean annual
temperature (MAT) and mean annual precipitation (MAP), as the most
commonly reported site-level climate variables. C fluxes increase
strongly with \(MAT\) on the global scale, but whether they saturate or
potentially decrease at higher temperatures remains disputed. Some
studies have detected no deceleration or decline in \(GPP\)
\citep{luyssaert_co_2007} or \(NPP\) \citep{schuur_productivity_2003}
with increasing MAT. Some have suggested that relationship between C
flux and temperature is exponential in nature \citep{allen_linking_2005}
(\textbf{REFS}). In contrast, others have found evidence of saturation
or decline of C flux in the warmest climates; \citet{luyssaert_co_2007}
found \(NPP\) saturating at around 10\(^\circ\)C MAT;
\citet{larjavaara_temperature_2012} found that increases in \(GPP\)
saturate at approximately 25\(^\circ\)C MAT, and
\citet{sullivan_sensitivity_2020} found that, within the tropics,
\(ANPP_{stem}\) decreases at the highest maximum temperatures. Various
measures of primary productivity saturate at high levels of MAP, though
the saturation points identified vary from 1500mm
\citep{luyssaert_co_2007} up to 2,445 mm MAP
\citep{schuur_productivity_2003}. Interactions between MAT and MAP are
also possible; within the tropics, there is a positive interaction
between MAT and MAP in shaping \(ANPP\), such that high rainfall has a
negative effect on productivity in cooler climates, compared to a
positive effect in warmer climates \citep{taylor_temperature_2017}.
\textbf{Such complicated dynamics play out with belowground fluxes as
well, with typically dominant soil temperature mediated by moisture and
C supply \citep{hursh_sensitivity_2017, xu_contribution_2016}.
\url{https://github.com/forc-db/Global_Productivity/issues/111}} There
is also evidence that C fluxes also respond to climate variables such as
temperature and precipitation seasonality (\textbf{REFS}), cloud cover
\citep{taylor_temperature_2017}, solar radiation
\citep{beer_terrestrial_2010, fyllas_solar_2017}, and potential
evapotranspiration \citep{kerkhoff_plant_2005}; however, these are not
typically assessed in global-scale analyses of annual forest C flux.

@@

We show that C fluxes are broadly consistent in their responses to
climate drivers on the global scale, with no trends in C allocation
among the variable pairs tested (Figs. 2, 4, S3). This parallels the
observation that C allocation across multiple C fluxes varies little
with respect to climate along a steep tropical elevational gradient
\citetext{\citealp{malhi_variation_2017}; \citealp[but
see][]{moser_elevation_2011}}, and is not surprising given that trees
face universal constraints in carbon allocation (\textbf{REFS}). We find
no trend in the allocation of \(GPP\) between production and respiration
across latitude or climate (\(NPP\):\(R_{auto}\); Fig. S3), refuting the
idea that tropical forests have anomalously low \(CUE\)
\citep{de_lucia_forest_2007, malhi_productivity_2012, anderson-teixeira_carbon_2016}.
Rather, differences in \(CUE\) between old-growth tropical forests
relative to (mostly younger) extratropical forests are likely an
artifact of comparing stands of different age, as \(CUE\) is known to
decline with forest age
\citep{de_lucia_forest_2007, piao_forest_2010, collalti_is_2019}.
Another previously observed pattern for which we find no support is a
tendency for belowground C allocation to decrease with increasing
temperature \citep{moser_elevation_2011, gill_belowground_2016}; rather,
we observe no trends in allocation between \(ANPP\) and \(BNPP\) across
latitudes. Failure to detect signficant tends in C allocation with
respect to climate in this analysis does not imply that none exist;
rather, it suggests that

Despite the broad consistency of climate responses across C fluxes,
climate explains lower proportions of variability among some of the
subsidiary C fluxes (\emph{e.g.}, \(ANPP_{stem}\), \(BNPP\)
\(BNPP_{fine.root)\); Fig. 2; Tables S2, S6). There are two,
non-exclusive, potential explanations for this. First, it may be that
methodological variation is larger relative to flux magnitude for some
of the subsidiary fluxes. Belowground fluxes in particular are difficult
to quantify, and measurement methods for the belowground fluxes
considered here may use fundamentally different approaches in different
sites (\emph{e.g.}, minirhizotrons, ingrowth cores, or sequential coring
for \(BNPP_{fine.root}\); root exclusion, stable isotope tracking, or
gas exchange of excised roots for \(R_{root}\)), and sampling depth is
variable and often insufficient to capture the full soil profile.
\(ANPP_{stem}\), which is also poorly explained by latitude or climate,
is more straightforward to measure but is subject to variability
introduced by differences such as biomass allometries applied and
minimum plant size sampled \citep{clark_measuring_2001}. However,
methodological variation and uncertainty affect all of fluxes considered
here, and some of the larger fluxes that vary more strongly with respect
to climate (\(ANPP\), \(NPP\)) are estimated by summing uncertain
component fluxes. Second, differences among variables in the proportion
of variation explained by climate may be attributable to more direct
climatic control over \(GPP\) than subsidiary fluxes. That is,
subsidiary fluxes may be shaped by climate both indirectly through its
influence on \(GPP\) and respiration and directly through any climatic
influence on C allocation, as well as many other local- and
regional-scale factors (\textbf{REFS}).

Temperature and its seasonality were the primary drivers of C flux on
the global scale (Figs. 2-4), consistent with a long legacy of research
identifying temperature as a primary driver of forest ecosystem C
cycling {[}e.g., \citet{lieth_primary_1973}; \textbf{REFS}{]}. We find
little evidence of any non-linearity in temperature's influence on C
flux. The relationship of all fluxes to \(MAT\) as an individual driver
were best described by a linear function (Table S2) - with the exception
of \(BNPP\), whose response to \(MAT\) was close to linear (Fig. 4a).
This result contrasts with both the idea that C flux responds
exponentially to temperature (\textbf{REFS}) and with the idea that
fluxes saturate with \(MAT\) below approximately 25\(^\circ\)C MAT
(\textbf{REFS}). It remains possible that flux declines above this
threshold
\citep{larjavaara_temperature_2012, sullivan_sensitivity_2020}, as is
also consistent with tree-ring records indicating that tropical tree
growth declines at high temperatures
\citep[e.g.,][]{vlam_temperature_2014}. However, these higher
temperatures also tend to be associated with high PET and VPD, both of
which are associated with reduced C fluxes (Figs. 4c-d, S4-S5). Indeed,
while temperature responses dominate at this global scale and with the
climate space occupied by forests, the effects of temperature are
moderated by moisture availability (Table 1; Figs 3-4). Specifically, C
fluxes are reduced under relatively dry conditions (\emph{i.e.}, low
\(MAP\); high \(VPD\)) and sometimes under very high precipitation
(Figs. 3-4). The observed positive interaction between \(MAT\) and
\(MAP\) for \(ANPP_{stem}\) on the global scale (Fig. 3) is consistent
with an analysis showing a similar interaction for \(ANPP\) in tropical
forests, also with a cross-over point at \textasciitilde20\(\circ\)C
\citep{taylor_temperature_2017}.\\
However, we detect no such interaction for \(ANPP\) or most other C
fluxes, and we find a contrasting negative interaction for \(NPP\) (Fig.
3), suggesting that more data are required to sort out potential
differences in the interactive effects of MAT and MAP on C fluxes in the
tropics.

@@

Our analysis clarifies how forest autotrophic carbon fluxes vary with
latitude and climate on a global scale, with some important implications
for how forest carbon cycling relates to climate and, by extension, how
it is likely to respond to climatic warming. To the extend that patterns
across broad scale climatic gradients can foretell how ecosystem
responses to climate change, our findings suggest that higher
temperatures with similar moisture availability result in a generalized
acceleration of forest C cycling (Figs. 2-3). This is consistent with
observations of continental- to global-scale increases over time in
\(GPP\) \citep{li_mapping_2019} and \(ANPP_{stem}\)
\citep{brienen_long-term_2015, hubau_asynchronous_2020}, along with some
C cycle components not considered here--tree mortality
\citep{brienen_long-term_2015, mcdowell_drivers_2018}, soil respiration
\citep{bond-lamberty_global_2010}, and heterotrophic soil respiration
\citep{bond-lamberty_globally_2018}. However, increasing C flux rates
are by no means universal
\citep[e.g.,][]{rutishauser_testing_2020, hubau_asynchronous_2020},
likely because other factors are at play, including changes to other
aspects of climate, atmospheric pollution (CO\textsubscript{2},
SO\textsubscript{2}, NO\textsubscript{x}), and local disturbances.\\
Moreover, forest ecosystem responses to climatic changes outside the
temperature range to which forest communities are adapted and
acclimatized will not necessarily parallel responses across geographic
gradients in climate. Indeed, tree-ring studies from forests around the
world indicate that tree growth rates - along with \(ANPP_{stem}\) and
possibly other ecosystem C fluxes - respond negatively to temperature
\citep{sniderhan_growth_2016, helcoski_growing_2019}. Furthermore, in
the tropics, climate change will push forests beyond any contemporary
climate, and there are some indications that this could reduce C flux
rates \citep{sullivan_sensitivity_2020} (\textbf{REFS}). Thus, while
further research is required to understand the extent to which forest
responses to climate change will track global gradients, and the time
scale on which they will do so, understanding the fundamental climatic
controls on annual C cycling in Earth's forests sets a firmer foundation
for understanding forest C cycle responses to accelerating climate
change.

@@

Scholarly Studies ForestGEO Compilation of the ForC database was
originally funded by DOE

>>>>

The latitudinal gradient in forest C flux rates, along with altitudinal
gradients \citep{girardin_net_2010, malhi_variation_2017}, is driven
primarily by climate, which is a significant driver of C fluxes across
broad spatial scales
\citep{luyssaert_co_2007, cleveland_relationships_2011, wei_forest_2010}.
However, there is little consensus as to the shapes of these
relationships or the best predictor variables. The majority of studies
have focused on exploring the relationships of C fluxes to mean annual
temperature (MAT) and mean annual precipitation (MAP), as the most
commonly reported site-level climate variables. C fluxes increase
strongly with \(MAT\) on the global scale, but whether they saturate or
potentially decrease at higher temperatures remains disputed. Some
studies have detected no deceleration or decline in \(GPP\)
\citep{luyssaert_co_2007}, \(NPP\) \citep{schuur_productivity_2003}, or
root respiration \citep[\(R_{root}\);][]{wei_forest_2010} with
increasing \(MAT\). In contrast, others have found evidence of
saturation or decline of C flux in the warmest climates;
\citet{luyssaert_co_2007} found \(NPP\) saturating at around
10\(^\circ\)C MAT; \citet{larjavaara_temperature_2012} found that
increases in \(GPP\) saturate at approximately 25\(^\circ\)C MAT, and
\citet{sullivan_long-term_2020} found that, within the tropics,
\(ANPP_{stem}\) decreases at the highest maximum temperatures. C fluxes
generally saturate at high levels of MAP, though the saturation points
identified vary from \(MAP\) of \textasciitilde1000 mm for \(R_{root}\)
\citep{wei_forest_2010} up to 2,445 mm for \(NPP\)
\citep{schuur_productivity_2003}. Interactions between MAT and MAP are
also possible; within the tropics, there is a positive interaction
between MAT and MAP in shaping \(ANPP\), such that high rainfall has a
negative effect on productivity in cooler climates, compared to a
positive effect in warmer climates \citep{taylor_temperature_2017}.
There is also evidence that C fluxes also respond to climate variables
such as temperature and precipitation seasonality
\citep{wagner_climate_2016}, cloud cover
\citep{taylor_temperature_2017}, solar radiation
\citep{beer_terrestrial_2010, fyllas_solar_2017}, and potential
evapotranspiration \citep{kerkhoff_plant_2005}; however, these are not
typically assessed in global-scale analyses of annual forest C flux.

@@

We show that C fluxes are broadly consistent in their responses to
climate drivers on the global scale, with no trends in C allocation
among the variable pairs tested (Figs. 2, 4, S3). This parallels the
observation that C allocation across multiple C fluxes varies little
with respect to climate along a steep tropical elevational gradient
\citetext{\citealp{malhi_variation_2017}; \citealp[but
see][]{moser_elevation_2011}}, and is not surprising given that carbon
allocation within forest ecosystems is relatively constrained
\citep{enquist_global_2002, litton_carbon_2007, malhi_allocation_2011}.
We find no trend in the allocation of \(GPP\) between production and
respiration across latitude or climate (\(NPP\):\(R_{auto}\); Fig. S3),
refuting the idea that tropical forests have anomalously low \(CUE\)
\citep{de_lucia_forest_2007, malhi_productivity_2012, anderson-teixeira_carbon_2016}.
Rather, differences in \(CUE\) between old-growth tropical forests
relative to (mostly younger) extratropical forests are likely an
artifact of comparing stands of different age, as \(CUE\) is known to
decline with forest age
\citep{de_lucia_forest_2007, piao_forest_2010, collalti_is_2019}.
Another previously observed pattern for which we find no support is a
tendency for belowground C allocation to decrease with increasing
temperature \citep{moser_elevation_2011, gill_belowground_2016}; rather,
we observe no trends in allocation between \(ANPP\) and \(BNPP\) across
latitudes. Failure to detect signficant tends in C allocation with
respect to climate in this analysis does not imply that none exist;
rather, it suggests that

Despite the broad consistency of climate responses across C fluxes,
climate explains lower proportions of variability among some of the
subsidiary C fluxes (\emph{e.g.}, \(ANPP_{stem}\), \(BNPP\)
\(BNPP_{fine.root}\); Fig. 2; Tables S2, S6). There are two,
non-exclusive, potential explanations for this. First, it may be that
methodological variation is larger relative to flux magnitude for some
of the subsidiary fluxes. Belowground fluxes in particular are difficult
to quantify, and measurement methods for the belowground fluxes
considered here may use fundamentally different approaches in different
sites (\emph{e.g.}, minirhizotrons, ingrowth cores, or sequential coring
for \(BNPP_{fine.root}\); root exclusion, stable isotope tracking, or
gas exchange of excised roots for \(R_{root}\)), and sampling depth is
variable and often insufficient to capture the full soil profile.
\(ANPP_{stem}\), which is also poorly explained by latitude or climate,
is more straightforward to measure but is subject to variability
introduced by differences such as biomass allometries applied and
minimum plant size sampled \citep{clark_measuring_2001}. However,
methodological variation and uncertainty affect all of fluxes considered
here, and some of the larger fluxes that vary more strongly with respect
to climate (\(ANPP\), \(NPP\)) are estimated by summing uncertain
component fluxes. Second, differences among variables in the proportion
of variation explained by climate may be attributable to more direct
climatic control over \(GPP\) than subsidiary fluxes. That is,
subsidiary fluxes may be shaped by climate both indirectly through its
influence on \(GPP\) and respiration and directly through any climatic
influence on C allocation, as well as many other local- and
regional-scale factors \citep[e.g.,][]{moser_elevation_2011}
(\textbf{REFS-BECKY, know any?}).

Temperature and its seasonality were the primary drivers of C flux on
the global scale (Figs. 2-4), consistent with a long legacy of research
identifying temperature as a primary driver of forest ecosystem C
cycling
\citep[e.g.,][]{lieth_primary_1973, luyssaert_co_2007, wei_forest_2010}.
We find little evidence of any non-linearity in temperature's influence
on C flux. The relationship of all fluxes to \(MAT\) as an individual
driver were best described by a linear function (Table S2) - with the
exception of \(BNPP\), whose response to \(MAT\) was close to linear
(Fig. 4a). This result contrasts with the idea that fluxes saturate with
\(MAT\) below approximately 25\(^\circ\)C \(MAT\)
\citep{luyssaert_co_2007, huston_global_2009}. It remains possible that
flux declines above this threshold
\citep{larjavaara_temperature_2012, sullivan_long-term_2020}, as is also
consistent with tree-ring records indicating that tropical tree growth
declines at high temperatures \citep[e.g.,][]{vlam_temperature_2014}.
However, these higher temperatures also tend to be associated with high
PET and VPD, both of which are associated with reduced C fluxes (Figs.
4c-d, S4-S5). Indeed, while temperature responses dominate at this
global scale and with the climate space occupied by forests, the effects
of temperature are moderated by moisture availability (Table 1; Figs
3-4). Specifically, C fluxes are reduced under relatively dry conditions
(\emph{i.e.}, low \(MAP\); high \(VPD\)) and sometimes under very high
precipitation (Figs. 3-4). The observed positive interaction between
\(MAT\) and \(MAP\) for \(ANPP_{stem}\) on the global scale (Fig. 3) is
consistent with an analysis showing a similar interaction for \(ANPP\)
in tropical forests, also with a cross-over point at
\textasciitilde20\(\circ\)C \citep{taylor_temperature_2017}.\\
However, we detect no such interaction for \(ANPP\) or most other C
fluxes, and we find a contrasting negative interaction for \(NPP\) (Fig.
3), suggesting that more data are required to sort out potential
differences in the interactive effects of MAT and MAP on C fluxes in the
tropics.

@@

Our analysis clarifies how forest autotrophic carbon fluxes vary with
latitude and climate on a global scale, with some important implications
for how forest carbon cycling relates to climate and, by extension, how
it is likely to respond to climatic warming. To the extend that patterns
across broad scale climatic gradients can foretell how ecosystem
responses to climate change, our findings suggest that higher
temperatures with similar moisture availability result in a generalized
acceleration of forest C cycling (Figs. 2-3). This is consistent with
observations of continental- to global-scale increases over time in
\(GPP\) \citep{li_mapping_2019} and \(ANPP_{stem}\)
\citep{brienen_long-term_2015, hubau_asynchronous_2020}, along with some
C cycle components not considered here--tree mortality
\citep{brienen_long-term_2015, mcdowell_drivers_2018}, soil respiration
\citep{bond-lamberty_global_2010}, and heterotrophic soil respiration
\citep{bond-lamberty_globally_2018}. However, increasing C flux rates
are by no means universal
\citep[e.g.,][]{rutishauser_testing_2020, hubau_asynchronous_2020},
likely because other factors are at play, including changes to other
aspects of climate, atmospheric pollution (CO\textsubscript{2},
SO\textsubscript{2}, NO\textsubscript{x}), and local disturbances.\\
Moreover, forest ecosystem responses to climatic changes outside the
temperature range to which forest communities are adapted and
acclimatized will not necessarily parallel responses across geographic
gradients in climate. Indeed, tree-ring studies from forests around the
world indicate that tree growth rates - along with \(ANPP_{stem}\) and
possibly other ecosystem C fluxes - respond negatively to temperature
\citep{sniderhan_growth_2016, helcoski_growing_2019}. Furthermore, in
the tropics, climate change will push forests beyond any contemporary
climate, and there are some indications that this could reduce C flux
rates \citep{mau_temperate_2018, sullivan_long-term_2020}. Thus, while
further research is required to understand the extent to which forest
responses to climate change will track global gradients, and the time
scale on which they will do so, understanding the fundamental climatic
controls on annual C cycling in Earth's forests sets a firmer foundation
for understanding forest C cycle responses to accelerating climate
change.

@@

We gratefully acknowledge all researchers who originally collected the
data used in this analysis. This study was funded by a Smithsonian
Scholarly Studies grant to KJAT and HCML and by Smithsonian's Forest
Global Earth Observatory (ForestGEO). Original compilation of the ForC
database was funded by \textbf{DOE grant to KAT}.
